---
title: "DroneBioR Biomass Report"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    df_print: paged
params:
  project_dir: !r getwd()
  field_csv: !r NULL
  use_alpha: true
  output_dir: !r tempfile("dronebior-report-")
  roi_geojson: !r NULL
  snapshot_path: !r NULL
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
project <- DroneBioR::dronebio_project(project_dir = params$project_dir)
suppressWarnings(DroneBioR::configure_proj_database(verbose = FALSE))
```

## Project

- **Directory:** `` `r project$project_dir` ``
- **Generated:** `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`
- **R session:** `r paste(R.version$major, R.version$minor, sep = ".")`
- **DroneBioR:** `r as.character(utils::packageVersion("DroneBioR"))`

### ODM / engine products

```{r products}
products <- DroneBioR::summarize_odm_products(project)
knitr::kable(products, caption = "Engine products discovered in the project directory.")
```

## Scientific workflow

```{r run-workflow}
result <- tryCatch(
  DroneBioR::run_dronebio_workflow(
    project    = project,
    use_alpha  = params$use_alpha,
    output_dir = params$output_dir
  ),
  error = function(e) {
    message("run_dronebio_workflow() failed: ", conditionMessage(e))
    NULL
  }
)
```

```{r workflow-status, results = "asis"}
if (is.null(result)) {
  cat("> The scientific workflow could not be executed. Spectral analysis sections are skipped.\n")
} else {
  cat("Workflow executed. Outputs written under: `", result$output_paths[["indices"]] |> dirname(), "`\n", sep = "")
}
```

### Reflectance summary

```{r refl-summary}
if (!is.null(result)) {
  knitr::kable(
    result$reflectance_summary,
    digits  = 4,
    caption = "Per-band reflectance summary."
  )
}
```

### Spectral index summary

```{r index-summary}
if (!is.null(result)) {
  knitr::kable(
    result$index_summary,
    digits  = 4,
    caption = "Per-index summary statistics."
  )
}
```

### Index histograms

```{r index-hist, fig.width = 9, fig.height = 7}
if (!is.null(result)) {
  layers <- names(result$indices)
  n <- length(layers)
  rows <- ceiling(n / 3)
  par(mfrow = c(rows, 3), mar = c(3, 3, 2, 1))
  for (lyr in layers) {
    vals <- terra::values(result$indices[[lyr]], mat = FALSE)
    vals <- vals[is.finite(vals)]
    if (length(vals) > 1) {
      hist(vals, main = lyr, xlab = "", col = "#9bdbcf", border = "white", breaks = 30)
    } else {
      plot.new(); title(main = paste(lyr, "(no data)"))
    }
  }
}
```

### Biomass proxy

```{r biomass-proxy, fig.width = 7, fig.height = 5}
if (!is.null(result) && !is.null(result$biomass_proxy)) {
  terra::plot(
    result$biomass_proxy,
    col  = hcl.colors(50, "YlGn"),
    main = "Biomass proxy (rescaled, image-only)"
  )
}
```

## Canopy height model

```{r chm, fig.width = 8, fig.height = 5}
dsm_path <- file.path(project$odm_project_dir, "odm_dem", "dsm.tif")
dtm_path <- file.path(project$odm_project_dir, "odm_dem", "dtm.tif")
chm <- tryCatch(
  DroneBioR::build_chm_from_dsm_dtm(dsm_path, dtm_path),
  error = function(e) NULL
)
if (!is.null(chm)) {
  terra::plot(
    chm,
    col  = hcl.colors(50, "viridis"),
    main = "Canopy Height Model (m)"
  )
}
```

```{r chm-summary}
if (!is.null(chm)) {
  vals <- terra::values(chm, mat = FALSE)
  vals <- vals[is.finite(vals) & vals > 0]
  if (length(vals) > 0) {
    knitr::kable(
      data.frame(
        metric = c("cells with height > 0", "min height (m)", "mean height (m)",
                   "max height (m)", "sd (m)"),
        value  = c(
          length(vals),
          formatC(min(vals), digits = 3, format = "f"),
          formatC(mean(vals), digits = 3, format = "f"),
          formatC(max(vals), digits = 3, format = "f"),
          formatC(stats::sd(vals), digits = 3, format = "f")
        )
      ),
      caption = "CHM summary."
    )
  }
}
```

## Field-based biomass model

```{r field-model, results = "asis"}
if (!is.null(params$field_csv) && nzchar(params$field_csv) &&
    file.exists(params$field_csv) && !is.null(result)) {
  field   <- DroneBioR::read_field_data(params$field_csv)
  joined  <- DroneBioR::extract_field_spectral_data(field, result$indices)
  model   <- DroneBioR::fit_biomass_lm(joined)
  r2      <- summary(model)$r.squared
  cat("**Predictors used:** ", paste(names(coef(model))[-1], collapse = ", "), "\n\n")
  cat("**Adjusted R^2:** ", round(summary(model)$adj.r.squared, 4), "\n\n")
  cat("**Multiple R^2:** ", round(r2, 4), "\n\n")
  print(knitr::kable(
    as.data.frame(summary(model)$coefficients),
    digits  = 4,
    caption = "Linear model coefficients."
  ))
} else {
  cat("> No field samples supplied (`field_csv` parameter). Add a CSV with ",
      "`sample_id`, `biomass_kgha` and x/y or longitude/latitude columns to ",
      "include a baseline biomass model in this report.\n", sep = "")
}
```

## 3D scene documentation

```{r 3d-snapshot, results = "asis"}
if (!is.null(params$snapshot_path) && nzchar(params$snapshot_path) &&
    file.exists(params$snapshot_path)) {
  cat("![3D scene snapshot from Drone Biomass Studio](", params$snapshot_path, ")\n\n", sep = "")
  cat("Snapshot captured from the 3D Modeling viewer using the Screenshot ",
      "toolbar button. The image is the WebGL framebuffer at the moment of ",
      "capture; HTML overlays (legend, scale bar, gizmo) are intentionally ",
      "excluded.\n", sep = "")
} else {
  cat("> No 3D viewer snapshot supplied. Falling back to a server-side ",
      "DSM perspective render below.\n\n", sep = "")
}
```

```{r 3d-persp, fig.width = 8, fig.height = 5}
if (is.null(params$snapshot_path) || !file.exists(params$snapshot_path %||% "")) {
  if (!is.null(chm)) {
    dsm <- tryCatch(terra::rast(dsm_path), error = function(e) NULL)
    if (!is.null(dsm)) {
      # Sub-sample to keep persp() responsive on big DSMs.
      ds <- terra::spatSample(dsm, size = 90000, method = "regular",
                              as.raster = TRUE, na.rm = FALSE)
      m  <- terra::as.matrix(ds, wide = TRUE)
      x  <- seq(terra::xmin(ds), terra::xmax(ds), length.out = ncol(m))
      y  <- seq(terra::ymin(ds), terra::ymax(ds), length.out = nrow(m))
      # `persp` wants z[i, j] where i indexes x and j indexes y. terra
      # returns rows = y (top down), so transpose and flip.
      z  <- t(m)[, rev(seq_len(nrow(m)))]
      par(mar = c(1, 1, 1, 1))
      persp(x, y, z,
            theta = 30, phi = 25, expand = 0.6,
            col   = "#bcd9b9",
            border = NA,
            shade = 0.55,
            ltheta = 135, lphi = 60,
            ticktype = "simple",
            xlab = "Easting (m)", ylab = "Northing (m)", zlab = "Z (m)")
    }
  }
}
```

## Survey-grade volumes per ROI

```{r survey-volumes, results = "asis"}
.format_num <- function(v, d = 3) {
  if (!is.finite(v)) "-" else formatC(v, format = "f", digits = d)
}
if (!is.null(params$roi_geojson) && nzchar(params$roi_geojson) &&
    file.exists(params$roi_geojson) && !is.null(result)) {
  rois <- tryCatch(sf::st_read(params$roi_geojson, quiet = TRUE),
                   error = function(e) NULL)
  if (is.null(rois) || nrow(rois) == 0) {
    cat("> ROI GeoJSON could not be read or is empty.\n")
  } else {
    dsm <- tryCatch(terra::rast(dsm_path), error = function(e) NULL)
    if (is.null(dsm)) {
      cat("> No DSM available for volume calculation.\n")
    } else {
      # Reproject ROIs to DSM CRS once.
      rois <- sf::st_transform(rois, terra::crs(dsm))
      methods <- c("dtm", "min_z", "mean_z", "perimeter_tin")
      out_rows <- list()
      for (i in seq_len(nrow(rois))) {
        nm <- if ("name" %in% names(rois)) as.character(rois$name[[i]]) else paste0("roi_", i)
        coords <- sf::st_coordinates(rois$geometry[[i]])
        roi_df <- data.frame(x = coords[, 1], y = coords[, 2])
        for (m in methods) {
          v <- tryCatch(
            DroneBioR::compute_survey_volumes(
              top    = dsm,
              roi    = roi_df,
              method = m,
              dtm    = if (identical(m, "dtm")) tryCatch(terra::rast(dtm_path), error = function(e) NULL) else NULL
            ),
            error = function(e) NULL
          )
          if (!is.null(v)) {
            out_rows[[length(out_rows) + 1L]] <- data.frame(
              roi            = nm,
              method         = m,
              cut_m3         = .format_num(v$cut_volume_m3),
              fill_m3        = .format_num(v$fill_volume_m3),
              net_m3         = .format_num(v$net_volume_m3),
              planar_m2      = .format_num(v$surface_area_planar_m2, 2),
              draped_m2      = .format_num(v$surface_area_draped_m2, 2),
              perimeter_m    = .format_num(v$perimeter_m, 2),
              cell_count     = as.character(v$cell_count),
              stringsAsFactors = FALSE
            )
          }
        }
      }
      if (length(out_rows) > 0) {
        print(knitr::kable(
          do.call(rbind, out_rows),
          caption = sprintf("Survey-grade volume metrics for %d ROI%s, four base-reference methods.",
                            nrow(rois), if (nrow(rois) == 1) "" else "s")
        ))
      }
    }
  }
} else {
  cat("> No ROIs supplied (parameter `roi_geojson`) or the file does not exist. ",
      "Draw ROIs in the GIS Workspace of Drone Biomass Studio and click ",
      "'Save ROI'; the report will pick up `studio_assets/rois.geojson` ",
      "automatically on the next render.\n", sep = "")
}
```

## About this report

This report was generated by `DroneBioR::render_dronebio_report()`. The
package wraps external photogrammetry engines (OpenDroneMap, WebODM,
Pix4Dmapper or Agisoft Metashape) and provides the scientific R layer:
radiometric scaling, vegetation indices, canopy height models, region
of interest analysis, and baseline biomass modelling.

Citation: run `citation("DroneBioR")` in R.
