---
title: "DroneBioR Biomass Report"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    df_print: paged
params:
  project_dir: !r getwd()
  field_csv: !r NULL
  use_alpha: true
  output_dir: !r tempfile("dronebior-report-")
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
project <- DroneBioR::dronebio_project(project_dir = params$project_dir)
suppressWarnings(DroneBioR::configure_proj_database(verbose = FALSE))
```

## Project

- **Directory:** `` `r project$project_dir` ``
- **Generated:** `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`
- **R session:** `r paste(R.version$major, R.version$minor, sep = ".")`
- **DroneBioR:** `r as.character(utils::packageVersion("DroneBioR"))`

### ODM / engine products

```{r products}
products <- DroneBioR::summarize_odm_products(project)
knitr::kable(products, caption = "Engine products discovered in the project directory.")
```

## Scientific workflow

```{r run-workflow}
result <- tryCatch(
  DroneBioR::run_dronebio_workflow(
    project    = project,
    use_alpha  = params$use_alpha,
    output_dir = params$output_dir
  ),
  error = function(e) {
    message("run_dronebio_workflow() failed: ", conditionMessage(e))
    NULL
  }
)
```

```{r workflow-status, results = "asis"}
if (is.null(result)) {
  cat("> The scientific workflow could not be executed. Spectral analysis sections are skipped.\n")
} else {
  cat("Workflow executed. Outputs written under: `", result$output_paths[["indices"]] |> dirname(), "`\n", sep = "")
}
```

### Reflectance summary

```{r refl-summary}
if (!is.null(result)) {
  knitr::kable(
    result$reflectance_summary,
    digits  = 4,
    caption = "Per-band reflectance summary."
  )
}
```

### Spectral index summary

```{r index-summary}
if (!is.null(result)) {
  knitr::kable(
    result$index_summary,
    digits  = 4,
    caption = "Per-index summary statistics."
  )
}
```

### Index histograms

```{r index-hist, fig.width = 9, fig.height = 7}
if (!is.null(result)) {
  layers <- names(result$indices)
  n <- length(layers)
  rows <- ceiling(n / 3)
  par(mfrow = c(rows, 3), mar = c(3, 3, 2, 1))
  for (lyr in layers) {
    vals <- terra::values(result$indices[[lyr]], mat = FALSE)
    vals <- vals[is.finite(vals)]
    if (length(vals) > 1) {
      hist(vals, main = lyr, xlab = "", col = "#9bdbcf", border = "white", breaks = 30)
    } else {
      plot.new(); title(main = paste(lyr, "(no data)"))
    }
  }
}
```

### Biomass proxy

```{r biomass-proxy, fig.width = 7, fig.height = 5}
if (!is.null(result) && !is.null(result$biomass_proxy)) {
  terra::plot(
    result$biomass_proxy,
    col  = hcl.colors(50, "YlGn"),
    main = "Biomass proxy (rescaled, image-only)"
  )
}
```

## Canopy height model

```{r chm, fig.width = 8, fig.height = 5}
dsm_path <- file.path(project$odm_project_dir, "odm_dem", "dsm.tif")
dtm_path <- file.path(project$odm_project_dir, "odm_dem", "dtm.tif")
chm <- tryCatch(
  DroneBioR::build_chm_from_dsm_dtm(dsm_path, dtm_path),
  error = function(e) NULL
)
if (!is.null(chm)) {
  terra::plot(
    chm,
    col  = hcl.colors(50, "viridis"),
    main = "Canopy Height Model (m)"
  )
}
```

```{r chm-summary}
if (!is.null(chm)) {
  vals <- terra::values(chm, mat = FALSE)
  vals <- vals[is.finite(vals) & vals > 0]
  if (length(vals) > 0) {
    knitr::kable(
      data.frame(
        metric = c("cells with height > 0", "min height (m)", "mean height (m)",
                   "max height (m)", "sd (m)"),
        value  = c(
          length(vals),
          formatC(min(vals), digits = 3, format = "f"),
          formatC(mean(vals), digits = 3, format = "f"),
          formatC(max(vals), digits = 3, format = "f"),
          formatC(stats::sd(vals), digits = 3, format = "f")
        )
      ),
      caption = "CHM summary."
    )
  }
}
```

## Field-based biomass model

```{r field-model, results = "asis"}
if (!is.null(params$field_csv) && nzchar(params$field_csv) &&
    file.exists(params$field_csv) && !is.null(result)) {
  field   <- DroneBioR::read_field_data(params$field_csv)
  joined  <- DroneBioR::extract_field_spectral_data(field, result$indices)
  model   <- DroneBioR::fit_biomass_lm(joined)
  r2      <- summary(model)$r.squared
  cat("**Predictors used:** ", paste(names(coef(model))[-1], collapse = ", "), "\n\n")
  cat("**Adjusted R^2:** ", round(summary(model)$adj.r.squared, 4), "\n\n")
  cat("**Multiple R^2:** ", round(r2, 4), "\n\n")
  print(knitr::kable(
    as.data.frame(summary(model)$coefficients),
    digits  = 4,
    caption = "Linear model coefficients."
  ))
} else {
  cat("> No field samples supplied (`field_csv` parameter). Add a CSV with ",
      "`sample_id`, `biomass_kgha` and x/y or longitude/latitude columns to ",
      "include a baseline biomass model in this report.\n", sep = "")
}
```

## About this report

This report was generated by `DroneBioR::render_dronebio_report()`. The
package wraps external photogrammetry engines (OpenDroneMap, WebODM,
Pix4Dmapper or Agisoft Metashape) and provides the scientific R layer:
radiometric scaling, vegetation indices, canopy height models, region
of interest analysis, and baseline biomass modelling.

Citation: run `citation("DroneBioR")` in R.
