---
title: "Reading products from ODM, WebODM, Pix4Dmapper and Agisoft Metashape"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading products from ODM, WebODM, Pix4Dmapper and Agisoft Metashape}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Why external engines

Photogrammetric reconstruction (SfM, MVS, mesh and texture generation) is
computationally heavy and is well served by established tools. DroneBioR does
not re-implement those algorithms in R — it consumes the products and runs
the scientific layer afterwards. The reader functions tolerate the small
layout differences between engines.

The minimum input DroneBioR needs is a **multispectral orthomosaic**: one
GeoTIFF with Blue, Green, Red, RedEdge and NIR layers and, optionally, an
alpha mask as a sixth layer. Optional additional inputs are a DSM, a DTM
and a dense point cloud.

## OpenDroneMap (ODM) via Docker

When ODM is installed locally, DroneBioR can drive it end-to-end.

```{r}
library(DroneBioR)

project <- dronebio_project("/path/to/Drone_Biomass")

# Stage MicaSense images into the ODM dataset folder
manifest <- list_micasense_images(project$images_dir)
copy_images_for_odm(manifest, project$odm_images_dir)

# Inspect the command that will be issued
build_odm_args(
  dataset_dir   = project$odm_dataset_dir,
  project_name  = project$odm_project_name,
  multispectral = TRUE
)

# Run ODM
run_odm_project(project, multispectral = TRUE)

# Summarize what ODM produced
summarize_odm_products(project)
```

The default ODM project layout is:

```
outputs/odm_micasense_dataset/
  micasense/
    images/                                          # input images
    odm_orthophoto/
      odm_orthophoto.tif                             # multispectral orthomosaic
    odm_dem/
      dsm.tif                                        # digital surface model
      dtm.tif                                        # digital terrain model
    odm_georeferencing/
      odm_georeferenced_model.laz                    # dense point cloud
```

`odm_product_paths(project)` returns these locations as a named list.

## WebODM

WebODM exports the same product set as ODM, usually as a downloadable
archive. Unpack it so the inner folder layout matches the ODM tree above,
then point `dronebio_project()` at the dataset directory. From that point
on, the workflow is identical to ODM.

## Pix4Dmapper

A typical Pix4Dmapper project lays its products out as:

```
3_dsm_ortho/
  2_mosaic/<name>_transparent_mosaic_group1.tif      # orthomosaic
  1_dsm/<name>_dsm.tif                               # DSM
2_densification/
  point_cloud/<name>_group1_densified_point_cloud.las
```

Point `read_multispectral_orthomosaic()` at the orthomosaic and pass the
DSM/DTM paths to `build_chm_from_dsm_dtm()`. Pix4D usually exports DTM in a
separate processing template — generate it once and reuse the file path.

```{r}
ortho <- read_multispectral_orthomosaic(
  "3_dsm_ortho/2_mosaic/flight_transparent_mosaic_group1.tif"
)
```

## Agisoft Metashape

Metashape writes products into an `export/` folder chosen at export time.
Recommended export settings:

* Orthomosaic: GeoTIFF, all five bands, optional alpha as a sixth layer.
* DSM and DTM: GeoTIFF, same CRS as the orthomosaic.
* Dense cloud: LAS or LAZ.

```{r}
ortho <- read_multispectral_orthomosaic("export/orthomosaic.tif")
chm   <- build_chm_from_dsm_dtm("export/dsm.tif", "export/dtm.tif")
```

## Reflectance scaling

Engine outputs differ in numeric range. MicaSense reflectance products are
typically 16-bit integers scaled by 32768. `scale_to_reflectance()`
auto-detects when values exceed 1; pass `scale_factor` explicitly when you
know the divisor your engine used.

```{r}
refl <- scale_to_reflectance(ortho$bands)         # auto-detects
refl <- scale_to_reflectance(ortho$bands, 32768)  # explicit
```

After scaling, `compute_spectral_indices(refl)` produces the index stack
regardless of which engine generated the orthomosaic. This is the boundary
where the scientific R layer takes over from the photogrammetry engine.
