---
title: "Dense point clouds, CHM and ROI metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dense point clouds, CHM and ROI metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Inputs

This vignette assumes you already have one or more of these products from
an external photogrammetry engine:

* a dense point cloud in LAS, LAZ or PLY;
* a DSM and a DTM as GeoTIFFs;
* a region of interest (a plot polygon, or a set of points to hull).

## Reading a point cloud

`read_full_point_cloud()` dispatches on file extension and returns a tidy
data frame with `x`, `y`, `z`, optional classification and a display color.

```{r}
library(DroneBioR)

pc <- read_full_point_cloud(
  "outputs/odm_micasense_dataset/micasense/odm_georeferencing/odm_georeferenced_model.laz"
)
nrow(pc)
```

For format-specific control:

```{r}
las <- read_las_point_cloud("dense.las")
ply <- read_ply_point_cloud("preview.ply", max_points = 50000)
```

The built-in LAS reader handles uncompressed `.las`. The PLY reader is
deliberately small for app previews — it expects the ODM/FPCFilter binary
little-endian layout (float x/y/z + uchar r/g/b/views). For very large LAZ
files, install the suggested `lidR` dependency, and `read_full_point_cloud()`
will use it transparently.

## CHM from DSM and DTM

```{r}
chm <- build_chm_from_dsm_dtm("export/dsm.tif", "export/dtm.tif")
```

When you only have a point cloud and no DTM, attach above-ground heights
using a low-quantile ground reference within each neighborhood:

```{r}
pc <- add_point_heights(pc, ground_quantile = 0.05)
```

When a CHM is already available, `add_chm_heights()` samples it at each
point so the heights are consistent with the raster product.

```{r}
pc <- add_chm_heights(pc, chm)
```

## Region-of-interest selection

A user selection is typically a convex hull around clicked points (in the
Shiny app) or a bounding box read from a field GIS. Both go through the
same polygon abstraction.

```{r}
roi <- build_roi_polygon(pc[sample(nrow(pc), 200), ], method = "hull")
# Or a bounding box:
roi_bbox <- build_roi_polygon(pc, method = "bbox")

inside <- points_in_roi(pc$x, pc$y, roi)
sel    <- filter_points_by_roi(pc, roi)
```

## CHM metrics for a ROI

```{r}
metrics <- compute_chm_roi_metrics(chm, roi)
metrics
```

Returned columns include area, mean and percentile heights, useful as
plot-level summaries to merge with field biomass measurements.

## Selection metrics and vertical profile

```{r}
sel <- add_point_heights(sel)

compute_selection_metrics(sel, voxel_size = 0.5)
compute_vertical_profile(sel, bin_size = 1)
```

`compute_selection_metrics()` returns a one-row data frame summarizing the
selection: n points, bounding-box area, max pairwise distance, height
percentiles and occupied voxel volume. `compute_vertical_profile()` returns
binned point counts useful for canopy structure plots.

## Individual tree candidates

```{r}
trees <- derive_tree_candidates(pc, min_height = 1.5)
```

The detector is intentionally simple — it produces seed locations for
downstream segmentation, not a final crown delineation.

## Exporting a selection

```{r}
export_point_selection(
  points     = sel,
  output_dir = "outputs/selections",
  base_name  = "plot_03"
)
```

Output files include a CSV table and a small PLY preview suitable for the
Shiny 3D viewer.

## Survey-grade volumes

`compute_survey_volumes()` computes the volume between a "top" surface
(a DSM, or any height raster) and a chosen base reference over a 2D
ROI. It is the function the Drone Biomass Studio app calls behind the
"Survey volumes" card in the 3D & Tree Metrics panel, and it can be
used directly from R for batch reporting.

Choose the base reference per use case:

| Method            | When to use it                                                                                                            |
|-------------------|---------------------------------------------------------------------------------------------------------------------------|
| `"dtm"`           | Canopy biomass when you have a separate DTM. Reduces to integrating CHM over the ROI - the standard biomass volume.       |
| `"min_z"`         | Classic stockpile - bare ground assumed to be at the lowest point inside the ROI.                                         |
| `"mean_z"`        | Mean-Z plane; good as a sanity check, but rarely the right answer.                                                        |
| `"ground_quantile"` | Robust low-Z plane (default 5th percentile). Useful when `"min_z"` is too sensitive to a single noisy pixel.            |
| `"user_plane"`    | Constant plane at a user-supplied Z. Use this when comparing flights against a fixed datum.                               |
| `"perimeter_tin"` | Industry standard for stockpile surveys (Pix4D Mapper, ContextCapture, Trimble Business Center). Builds a Delaunay TIN through the perimeter vertices and integrates `top - TIN` inside the ROI. Requires the optional `interp` package. |

```{r}
dsm <- terra::rast(system.file("extdata", "dsm_subset.tif", package = "DroneBioR"))
dtm <- terra::rast(system.file("extdata", "dtm_subset.tif", package = "DroneBioR"))
roi <- data.frame(
  x = c(392004, 392012, 392012, 392004),
  y = c(3033004, 3033004, 3033012, 3033012)
)
canopy <- compute_survey_volumes(top = dsm, roi = roi, method = "dtm", dtm = dtm)
stockpile <- compute_survey_volumes(top = dsm, roi = roi, method = "perimeter_tin")
canopy$cut_volume_m3
stockpile$cut_volume_m3
```

The result is a `list` with class `"dronebio_survey_volume"`. Print it
to see all the reported quantities (cut / fill / net volumes,
planimetric and 3D draped surface areas, perimeter, cell area and
count, and per-surface Z summaries).

## When to reach for `lidR`

For full LAZ analytics — tiled processing, advanced ground classification,
DTM modelling, terrain normalization, individual tree segmentation — the
`lidR` package is the right tool. DroneBioR's point-cloud functions are
deliberately scoped to the in-memory analyses that drive the Shiny app and
plot-level summaries.
