---
title: "Vegetation indices in DroneBioR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vegetation indices in DroneBioR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Reflectance is required

All indices in DroneBioR assume that input bands are scaled to reflectance
in the 0 – 1 range. MicaSense RedEdge orthomosaics are typically stored as
16-bit integers scaled by 32768; route them through `scale_to_reflectance()`
before computing indices.

```{r}
library(DroneBioR)

ortho <- read_multispectral_orthomosaic("path/to/orthomosaic.tif")
refl  <- scale_to_reflectance(ortho$bands)
```

The named layers in `refl` must be `Blue`, `Green`, `Red`, `RedEdge` and
`NIR`. `read_multispectral_orthomosaic()` enforces this naming when the band
map is correct.

## The index catalogue

`compute_spectral_indices()` returns a `terra::SpatRaster` with the layers
listed below, all computed with a small-denominator guard (`eps = 1e-6`)
to avoid division blow-ups in shadow or water pixels.

| Layer       | Formula                                                  | Typical use                                              |
|-------------|----------------------------------------------------------|----------------------------------------------------------|
| NDVI        | (NIR - Red) / (NIR + Red)                                | Greenness, general vigor; saturates at high LAI          |
| NDRE        | (NIR - RedEdge) / (NIR + RedEdge)                        | Chlorophyll / nitrogen status in dense canopies          |
| EVI         | 2.5 * (NIR - Red) / (NIR + 6*Red - 7.5*Blue + 1)         | High-biomass canopies, atmosphere-resistant              |
| SAVI        | (1 + L) * (NIR - Red) / (NIR + Red + L), L = 0.5         | Sparse canopies where soil reflectance is visible        |
| NDWI        | (Green - NIR) / (Green + NIR)                            | Water / canopy water content; flags wet pixels           |
| GNDVI       | (NIR - Green) / (NIR + Green)                            | Chlorophyll, sensitive at low-to-mid biomass             |
| CIrededge   | (NIR / RedEdge) - 1                                      | Chlorophyll without saturation at high LAI               |
| MSAVI2      | 0.5 * (2*NIR + 1 - sqrt((2*NIR + 1)^2 - 8*(NIR - Red)))  | Soil-adjusted, no tuned L parameter                      |
| VARI        | (Green - Red) / (Green + Red - Blue)                     | RGB-only proxy when NIR is unreliable                    |

## Computing the stack

```{r}
indices <- compute_spectral_indices(refl)
names(indices)
#> [1] "NDVI" "NDRE" "EVI" "SAVI" "NDWI" "GNDVI" "CIrededge" "MSAVI2" "VARI"

summarize_spatraster(indices)
```

## Biomass proxy

`compute_biomass_proxy()` blends a small subset of indices into a single
rescaled layer useful for quick visual screening. It is a proxy, not a
calibrated estimate — always validate against field samples before reporting
biomass values.

```{r}
proxy <- compute_biomass_proxy(indices)
```

## Choosing predictors for a biomass model

`fit_biomass_lm()` defaults to whichever of `NDVI`, `NDRE`, `EVI`, `SAVI`,
`NDWI`, `NIR` and `RedEdge` are present in the joined field/predictor table.
Pass `predictors = ...` to constrain the model — for dense canopies,
`c("NDRE", "CIrededge")` often outperforms NDVI alone.

```{r}
joined <- extract_field_spectral_data(field, indices)
model  <- fit_biomass_lm(joined, predictors = c("NDRE", "CIrededge"))
summary(model)
```

## Practical tips

* Always inspect the indices for hot pixels at the orthomosaic edges — the
  alpha mask in `read_multispectral_orthomosaic(..., use_alpha = TRUE)`
  removes most of them.
* When comparing flights, scale to reflectance first; otherwise NDVI shifts
  across dates reflect the engine's scaling, not the canopy.
* Index choice should follow the canopy stage. NDVI plateaus once LAI is
  above ~3; CIrededge and NDRE keep responding.

## Color stretch in the Shiny app

The GIS Workspace map offers three colour-stretch modes via the
"Color stretch" selector in the sidebar. The default is **Fixed semantic**
for a reason worth knowing if you publish or report from these maps.

* **Fixed semantic** (default). The colour scale is pinned to a canonical
  range per index: `[-1, 1]` for NDVI, NDRE, NDWI, GNDVI, VARI and the
  biomass proxy; `[0, 1]` for MSAVI2 and raw reflectance bands; the data
  range for EVI, SAVI and CIrededge (which can exceed those bounds in
  practice). With the Red-Yellow-Green palette, **yellow falls at zero** —
  marking the biophysical boundary between non-vegetation pixels (water,
  bare soil, shadow, asphalt) and vegetation. Identical index values
  always produce identical colours, which is what makes a time-series
  panel like the one in the *Time Series* tab actually comparable across
  dates.

* **Data range**. Linearly stretched between the layer's own minimum and
  maximum. Useful for highlighting subtle spatial structure within one
  scene, but the **same value will get different colours** in different
  flights if their min/max differs — so this mode is *not* appropriate
  when you compare multiple dates or sites.

* **Percentile 2-98**. Robust stretch that ignores the tails. Helpful
  when a few outlier pixels (sun glint, hot edge artefacts) compress the
  visible contrast under "Data range".

In remote-sensing publications, fixed semantic ranges are the convention
for diverging vegetation indices: the colour carries physical meaning
(green = vegetated, red = non-vegetated), and the reader can compare two
maps side by side. We keep that as the default; per-scene stretching is
available when you need it for exploration but is *not* the recommended
choice for the figures you ship in a paper.
